Contribution Title\thanks{Supported by organization x.}

### 2021-02-02 papers/actus2021/article.tex
more of actus article (#102)

%% LyX 2.3.6.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[runningheads]{llncs}
\usepackage[latin9]{inputenc}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}

@@

\begin{document}
\title{Contribution Title\thanks{Supported by organization x.}}
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and Second
Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
\authorrunning{F. Author et al.}
\institute{Princeton University, Princeton NJ 08544, USA \and Springer Heidelberg,
Tiergartenstr. 17, 69121 Heidelberg, Germany }
\email{lncs@springer.com}
\institute{\url{http://www.springer.com/gp/computer-science/lncs} \and ABC
Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
 }
\email{\{abc,lncs\}@uni-heidelberg.de}
\maketitle
\begin{abstract}
Crypto-loans are unique and innovative financial instruments that
allow trustless P2P lending, which could provide a safe and convenient
source of liquidity for cryptocurrency holders. In this article, we
explore a smart-contract framework for building standardized crypto-loans
with Marlowe DSL and Algorithmic Contract Types Unified Standards
at its core.

@@

\begin{lstlisting}
--Definitions/ContractState.hs 
data ContractStatePoly a b = ContractStatePoly  
{  
tmd     :: b  
, nt    :: a  
, ipnr  :: a  
, ipac  :: a  
, feac  :: a  
, fac   :: a  
, nsc   :: a  
, isc   :: a  
, prf   :: ContractStatus  
, sd    :: b  
, prnxt :: a  
, ipcb  :: a  
} deriving (Show)

--Ops.hs
class ActusOps a where    
   _min :: a -> a -> a
   _max :: a -> a -> a
   _zero :: a
   _one :: a

@@

Thus, every formula in the executable spec could either be instantiated
to: 
\begin{itemize}
\item Formula on some ``atomic'' type (like Double, Day) - could be used
to directly compute cash-flows for analytical purposes or precompute
payoffs for smart contracts that don't depend on oracles 
\item Formula representing piece of AST (Marlowe Value, Observation) - could
be used to generate smart contracts that depend on oracles or to generate
code in another language (like Agda) 
\end{itemize}
This approach of abstracting formulas has a limitation of not allowing
conditionals to be expressed in an abstract way (There is no ActusIf
typeclass) - luckily most of conditional expressions in ACTUS specification
don't depend on variable state of a contract - they depend on ContractTerms
that are known in advance during contract generation. This allows
us to dispatch appropriate formulas during generation rather than
execution.

@@

In order to simplify serialisation/deserialisation of contract terms
across Actus related services maintained by Cardano/IOG - we rely
on \textquotedblleft superposed\textquotedblright{} representation
of contract terms - all contract types are represented with the same
type. Although it allows both encoder and decoder to express any ACTUS
contract terms, it also allows for invalid combinations of terms (for
example PRNXT cannot be applied to PAM-contract), which means contracts
require specific validation that is implemented by means of applicability
function:

@@

Actus has a family of applicability functions polymorphic on contract
type:

\noindent 
\begin{equation}
Applicability:ContractType\times ContractTerm\mapsto ApplicabilityType
\end{equation}

Where applicability could be: none, always, nullable, multiple

@@

@@

\noindent 
\begin{equation}
contract(ct)=receiveCollaterals(ct)\circ INIT(ct)\circ\prod_{t\in SCHED(ct)}chainlink(t)
\end{equation}

where $receiveData$ - is asking an Oracle for Marlowe Choice if needed,
$calculatePayoff$ - calculates payoff formula. For fixed-rate contracts
this is optimised into a precalculated constant. process\_payoff -
await Deposit of payoff amount from a party: if deposit is transferred
it redirects the funds to a counterparty - otherwise it transfers
the collateral to a counterparty and closes the contract. The chain
is generated from the fixed (known in advance) schedule of events
using $SCHED()$function from ACTUS.

@@

Marlowe doesn\textquoteright t allow functions and recursion. Naive
usage of If operator in Marlowe could lead to exponential growth of
a contract, for example:

\begin{lstlisting}
If (condition) 
	then 
		perform_something1()
		continue() 
	else 
		perform_something2 
		continue()
\end{lstlisting}

would inline contents of \textquotedblleft continue()\textquotedblright{}
twice and given that ACTUS contracts are essentially generated using
continuation as an accumulator - this would lead to exponential explosion
of the size of any actus contract that has conditionals in their state
transition logic.

@@

adjusted = max(min(original, floor), cap) Where max(a, b) = if (a
\textgreater{} b) then a else b min(a,b) = if (a \textless{} b) then
a else b

We addressed this issue by introducing Cond(link to Marlowe docs)
construct in order to represent \textquotedblleft if\textquotedblright{}
expressions. So instead of using the \textquotedblleft If\textquotedblright{}
contract to decide the value of some variable - we use conditional
assignment instead as Cond is essentially just a pure function that
returns value depending on condition in contrast to the \textquotedblleft If\textquotedblright{}
contract that performs action instead of returning value. The name
is borrowed from a similar function in Lisp.

@@

Marlowe doesn\textquoteright t allow perpetual contracts even if their
recursion is productive (example: perpetual swap) - thus we can\textquoteright t
support certain contract types from ACTUS specification, namely the
ones that don\textquoteright t have a defined maturity date (like
UMP). In a long-term perspective - there is a possible workaround:
contracts with no maturity date could be represented as actors with
finite amounts of state transitions. We prototyped this approach in
a separate branch, however it does seem more prone to errors comparing
to rendering predefined schedules and more importantly it greatly
affects static analysis as the amount of \textquotedblleft reduction
steps\textquotedblright{} grows from N\_scheduled\_event = count(event)
to N\_state\_transition = max(date(event)) - min(date(event)) / precision.
See https://github.com/dk14/plutus/tree/ACTUS

@@

Marlowe only supports Integers, while ACTUS is expressed in terms
of real numbers. In order to model a real number in such a setup -
we rely on fixed point precision. The algebra looks like this: 
\begin{lstlisting}
   (+)                         = AddValue
   (-)                         = SubValue
   a * b                       = Scale (1 % marloweFixedPoint) $ MulValue a b
   (Constant 0) / (Constant 0) = (Constant 0)
   (Constant x) / (Constant y) = Scale (marloweFixedPoint % 1) $ Constant $ div x y
   x / (Constant y)            = Scale (marloweFixedPoint % y) $ x    
   _ / _                       = undefined --division not supported in Marlowe yet 
\end{lstlisting}

@@

Unfortunately, Marlowe DSL doesn\textquoteright t support any notion
of records, the variables can only be of type \textquotedblleft Integer\textquotedblright .
Thus, in order to map contract state (ContractStatePoly) we first
pack a set of Marlowe variables (Value Observation) representing previous
t-1 state into ContractStatePoly, then we apply polymorphic state
transition, and finally unpack ContractState poly into a set of Marlowe
variables into t+1 state:

st\_t+1 = unpack(stateTransition(pack(st\_t-1))) , where pack is a
chain of UseValue and unpack is a chain of Let

Representing \textquotedblleft mutations\textquotedblright{} (e.g.
state transitions) in Marlowe (unfolding)

Marlowe doesn\textquoteright t support mutable variables, so we have
to represent state at t (st\_t) literally by naming convention:

variable\_name(name, t) = concat(name, \textquotedblleft \_\textquotedblright ,
t) generate\_variable\_accessor(name, t) = UseValue variable\_name(name,
t) generate\_variable\_setter(name, t, formula) = Let variable\_name(name,
t) formula

@@

In order to demonstrate and test the capabilities of Actus generators,
a visual Blockly-based tool was developed for Marlowe Playground.
The Actus Labs tool allows users to construct contract terms visually,
generate a corresponding Marlowe contract and try it out in a simulated
environment.

@@

Every participant of an ACTUS contract is described by a Role which
is in its turn represented through a special unique non-fungible token.
This makes every Marlowe contract a tradable security, allowing a
participant to sell its \textquotedblleft share\textquotedblright{}
in a contract by selling a corresponding role token.

@@

For example, if Alice has fixed income from a loan (or some other
investment) and Bob has comparable but variable (fluctuating) income
- Bob can hedge by swapping cash flows with Alice. If Alice\textquoteright s
income is locked with token1 and Bob\textquoteright s income is locked
with token2 - then atomic swap of those tokens is equivalent to a
swap of cash flows.

@@

Forall ct. get\_cashflows\_haskell(ct, riskfactormodel) == get\_cashflows\_java(ct,
riskfactormodel)

Where get\_cashflows returns a set of \{date, payoff\} tuples.

@@

QuickCheck contract terms generators also allow to ensure arbitrary
properties of a Marlowe contract. This could be enhanced with Marlowe\textquoteright s
static analysis feature by utilising Assert operator:

ContractTerms := sample(qcgenerator) Contract := generateMarloweContract(ContractTerms)
ContractWithAssert := appendAssertion(Contract, assertion) runStaticAnalysis(ContractWithAssert)

While it doesn\textquoteright t cover all possible contracts it could
guarantee that property holds for a statistically significant fraction
of a contract. Using Static Analysis to check a particular contract
instead of randomly sampling allows: 
\begin{itemize}
\item more balanced sampling: Space of contract terms depends linearly on
the coverage, e.g. if we cover 10\% of all possible contract terms
- we\textquoteright ll cover 10\% of all contracts. Different contracts
have different amount of risk factors thus different search space,
for instance a contract with 3 risk factors (e.g. 3 observations of
an interest rate) would span over n\textasciicircum 3 while a contract
with 10 risk factors would span over n\textasciicircum 10. Meanwhile,
space of all risk factors in all contracts doesn\textquoteright t
linearly depend on the coverage. 
\item dependency tracking: SMT-solver would likely be more aware of execution
paths leading to failure of a test, which could significantly reduce
search space 
\item completeness: If not timed out, SMT-solver is decidable while sampling
is semi-decidable. 
\end{itemize}

@@

By design, the Marlowe interpreter always refunds assets held in a
contract when it terminates. This is done in order to ensure that
no funds are lost forever. The funds are returned to whoever\textquoteright s
internal account actually holds them.

@@

This implicit refund is easy to overlook by smart-contract developers
as plain Close is often used as a default action in case of unexpected
behaviour like timeouts(especially Choice timeout). This can easily
lead to multi-million dollar mistakes if malicious Alice decides to
exploit an auto-refund feature in order to get her collateral without
paying back. Here\textquoteright s the example of simple attack vector:
\begin{enumerate}
\item Alice creates a contract where Deposit timeout would lead to Close 
\item Bob doesn\textquoteright t know or test the timeout path. Even if
Bob is a programmer - he might not be aware that Closing contract
would cause collateral refund 
\item Alice puts collateral and gets notional from Bob 
\item Alice doesn\textquoteright t pay for a loan (Deposit timeout) 
\item Alice gets her collateral back 
\item Bob loses notional
\end{enumerate}
In order to prevent this from happening, an additional \textquotedblright Auto-Refund\textquotedblright{}
security check was introduced as part of Static Analysis {[}TODO link{]}
tooling which notifies users about all Close constructs that can lead
to automatic refunds and encourages users to write explicit logic
for edge cases.

@@

Unlike current popular DeFi lending approaches (like Harvest Finance)
- Marlowe ACTUS-based implementation tends to rely on \textquotedblleft good
old\textquotedblright{} matching instead of pooling. While pooling
is proven to be superior to order-book based approaches when it comes
to automated market makers, using it for lending has been shown to
be more susceptible to attacks (for example \textquotedblleft flash
loan\textquotedblright{} (https://www.coindesk.com/harvest-finance-24m-attack-triggers-570m-bank-run-in-latest-defi-exploit)).

@@

\begin{figure}
\includegraphics[width=1\textwidth]{fig1} \caption{A figure caption is always placed below the illustration. Please note
that short captions are centered, while long ones are justified by
the macro package automatically.}
\label{fig1} 
\end{figure}

>>>>

%% LyX 2.3.6.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[runningheads]{llncs}
\usepackage[latin9]{inputenc}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[unicode=true,
 bookmarks=false,
 breaklinks=false,pdfborder={0 0 1},backref=section,colorlinks=false]
 {hyperref}

@@

\begin{document}
\title{Contribution Title\thanks{Supported by organization x.}}

\maketitle
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here

\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and Second
Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
\authorrunning{F. Author et al.}

% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.

\institute{Princeton University, Princeton NJ 08544, USA \and Springer Heidelberg,
Tiergartenstr. 17, 69121 Heidelberg, Germany }
\email{lncs@springer.com}
\institute{\\
 \url{http://www.springer.com/gp/computer-science/lncs} \and ABC
Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
 }
\email{\{abc,lncs\}@uni-heidelberg.de}
\begin{abstract}
Crypto-loans are unique and innovative financial instruments that
allow trustless P2P lending, which could provide a safe and convenient
source of liquidity for cryptocurrency holders. In this article, we
explore a smart-contract framework for building standardized crypto-loans
with Marlowe DSL and Algorithmic Contract Types Unified Standards
at its core.

@@

\begin{lstlisting}
-- Definitions/ContractState.hs 
data ContractStatePoly a b = ContractStatePoly  
{  
tmd     :: b  
, nt    :: a  
, ipnr  :: a  
, ipac  :: a  
, feac  :: a  
, fac   :: a  
, nsc   :: a  
, isc   :: a  
, prf   :: ContractStatus  
, sd    :: b  
, prnxt :: a  
, ipcb  :: a  
} deriving (Show)

-- Ops.hs
class ActusOps a where    
   _min :: a -> a -> a
   _max :: a -> a -> a
   _zero :: a
   _one :: a

@@

Thus, every formula in the executable spec could either be instantiated
to: 
\begin{itemize}
\item formula on some ``atomic'' type (like Double, Day) - could be used
to directly compute cash-flows for analytical purposes or precompute
payoffs for smart contracts that don't depend on oracles 
\item formula representing piece of AST (Marlowe Value, Observation) - could
be used to generate smart contracts that depend on oracles or to generate
code in another language (like Agda) 
\end{itemize}
This approach of abstracting formulas has a limitation of not allowing
conditionals to be expressed in an abstract way (There is no ActusIf
typeclass) - luckily most of conditional expressions in ACTUS specification
don't depend on variable state of a contract - they depend on ContractTerms
that are known in advance during contract generation. This allows
us to dispatch appropriate formulas during generation rather than
execution.

@@

In order to simplify serialisation/deserialisation of contract terms
across Actus related services maintained by Cardano - we rely on ``superposed''
representation of contract terms - all ACTUS contract types are represented
with the same type. While such representation allows both encoder
and decoder to express any ACTUS contract terms, it also allows for
invalid combinations of terms (for example PRNXT cannot be applied
to PAM-contract), which means contracts require specific validation
that is implemented by means of applicability function:

@@

ACTUS defines a family of applicability functions polymorphic on contract
type (TODO link to ACTUS applicability):

\noindent 
\begin{equation}
Applicability:ContractType\times ContractTerm\mapsto ApplicabilityType,
\end{equation}

where applicability could be: none, always, nullable, multiple

@@

\begin{figure}
\includegraphics[width=1\textwidth]{modules} \caption{Modules responsible for contract generation.}
\label{fig1-1} 
\end{figure}

@@

\noindent 
\begin{equation}
contract(ct)=collaterals(ct)\circ INIT(ct)\circ\prod_{t\in SCHED(ct)}chainlink(t)
\end{equation}

where $receiveData$ - is asking an Oracle for Marlowe Choice if needed,
$calculatePayoff$ - calculates payoff formula. For fixed-rate contracts
this is optimised into a precalculated constant. $processPayoff$
- awaits Deposit of payoff amount from a party: if deposit is transferred
it redirects the funds to a counterparty - otherwise it transfers
the collateral to a counterparty and closes the contract. The chain
is generated from the fixed (known in advance) schedule of events
using $SCHED()$function from ACTUS.

\begin{figure}
\includegraphics[width=1\textwidth]{flowchart} \caption{Modules responsible for contract generation.}
\label{fig1-1-1} 
\end{figure}

@@

Marlowe doesn't allow functions and recursion. Naive usage of If operator
in Marlowe could lead to exponential growth of a contract, for example:

\begin{lstlisting}
if condition 
	then 
		perform_something1()
		continue() 
	else 
		perform_something2 
		continue()
\end{lstlisting}

would inline contents of $continue()$ twice and given that ACTUS
contracts are essentially generated using continuation as an accumulator
- this would lead to exponential explosion of the size of any actus
contract that has conditionals in their state transition logic.

@@

\noindent 
\begin{equation}
adjusted=max(min(original,floor),cap)
\end{equation}

We addressed this issue by introducing ``Cond'' construct in order
to represent ``if'' expressions. So instead of using the ``If''
contract to decide the value of some variable - we use conditional
assignment instead as ``Cond'' is essentially just a pure function
that returns value depending on condition in contrast to the ``If''
contract that performs action instead of returning value. The name
is borrowed from a similar function in Lisp.

@@

Marlowe doesn't allow perpetual contracts even if their recursion
is productive (example: perpetual swap) - thus we can't support certain
contract types from ACTUS specification, namely the ones that don't
have a defined maturity date (like UMP). In a long-term perspective
- there is a possible workaround: contracts with no maturity date
could be represented as actors with finite amounts of state transitions.
We prototyped this approach in a separate branch, however it does
seem more prone to errors comparing to rendering predefined schedules
and more importantly it greatly affects static analysis as the amount
of ``reduction steps'' grows from $N_{scheduled}=count(event)$
to $N_{stateTransitions}=\frac{max(date(event))-min(date(event))}{precision}$.
See \href{https://github.com/dk14/plutus/tree/ACTUS}{https://github.com/dk14/plutus/tree/ACTUS}

@@

Marlowe only supports Integers, while ACTUS is expressed in terms
of real numbers. In order to model a real number in such a setup -
we rely on fixed point precision. The algebra looks like this: 
\begin{lstlisting}
(+)   = AddValue
(-)   = SubValue
a * b = Scale (1 % marloweFixedPoint) $ MulValue a b
\end{lstlisting}

@@

Unfortunately, Marlowe DSL doesn't support any notion of records,
the variables can only be of type ``Integer''. Thus, in order to
map contract state (ContractStatePoly) we first pack a set of Marlowe
variables (Value Observation) representing previous $t-1$ state into
ContractStatePoly, then we apply polymorphic state transition, and
finally unpack ContractState poly into a set of Marlowe variables
into $t+1$ state:

\noindent 
\begin{equation}
st_{t+1}=unpack(stateTransition(pack(st_{t-1}))),
\end{equation}

,where pack is a chain of ``UseValue'' and unpack is a chain of
``Let''

\paragraph{Representing state transitions in Marlowe (unfolding)}

Marlowe doesn't support mutable variables, so we have to represent
state at t ($st_{t}$) literally through naming convention:

\begin{lstlisting}
variableName(name, t) = 
	concat(name, '_', t) 
generateAccessor(name, t) = 
	UseValue variableName(name, t) 
generateSetter(name, t, formula) = 
	Let variableName(name, t) formula 
\end{lstlisting}

@@

\begin{figure} 
\includegraphics[width=0.8\textwidth,natwidth=664,natheight=552]{labs.png}
\caption{Actus Labs - online tool for generating Actus contracts for Marlowe.} 
\label{fig2} 
\end{figure}

In order to demonstrate and test the capabilities of Actus generators,
a visual online Blockly-based tool was developed for Marlowe Playground.
The Actus Labs tool allows users to construct contract terms visually,
generate a corresponding Marlowe contract and try it out in a simulated
environment.

@@

Every participant of an ACTUS contract is described by a Role which
is in its turn represented through a special unique non-fungible token.
This makes every Marlowe contract a tradable security, allowing a
participant to sell its ``share'' in a contract by selling a corresponding
role token.

@@

For example, if Alice has fixed income from a loan (or some other
investment) and Bob has comparable but variable (fluctuating) income
- Bob can hedge by swapping cash flows with Alice. If Alice's income
is locked with token1 and Bob's income is locked with token2 - then
atomic swap of those tokens is equivalent to a swap of cash flows.

@@

\noindent 
\begin{equation}
\forall ct\forall rf.getCashFlows("haskell",ct,rf)\equiv getCashFlows("java",ct,rf)
\end{equation}

Where $ct$ represnts contract terms, $rf$ is risk factor model and
$getCashFlows()$ returns a set of $(date,payoff)$ tuples.

@@

QuickCheck contract terms generators also allow to ensure arbitrary
properties of a Marlowe contract. This could be enhanced with Marlowe's
static analysis feature by utilising Assert operator:

\begin{lstlisting}
ContractTerms := sample(qcgenerator) 
Contract := generateMarloweContract(ContractTerms) 
ContractWithAssert := appendAssertion(Contract, assertion) 
runStaticAnalysis(ContractWithAssert)
\end{lstlisting}

While it doesn't cover all possible contracts it could guarantee that
property holds for a statistically significant fraction of a contract.
Using Static Analysis to check a particular contract instead of randomly
sampling allows: 
\begin{itemize}
\item more balanced sampling: Space of contract terms depends linearly on
the coverage, e.g. if we cover 10\% of all possible contract terms
- we'll cover 10\% of all contracts. Different contracts have different
amount of risk factors thus different search space, for instance a
contract with 3 risk factors (e.g. 3 observations of an interest rate)
would span over n\textasciicircum 3 while a contract with 10 risk
factors would span over n\textasciicircum 10. Meanwhile, space of
all risk factors in all contracts doesn't linearly depend on the coverage. 
\item dependency tracking: SMT-solver would likely be more aware of execution
paths leading to failure of a test, which could significantly reduce
search space 
\item completeness: If not timed out, SMT-solver is decidable while sampling
is semi-decidable. 
\end{itemize}

@@

By design, the Marlowe interpreter always refunds assets held in a
contract when it terminates. This is done in order to ensure that
no funds are lost forever. The funds are returned to whoever's internal
account actually holds them.

@@

This implicit refund is easy to overlook by smart-contract developers
as plain Close is often used as a default action in case of unexpected
behaviour like timeouts(especially Choice timeout). This can easily
lead to multi-million dollar mistakes if malicious Alice decides to
exploit an auto-refund feature in order to get her collateral without
paying back. Here's the example of simple attack vector: 
\begin{enumerate}
\item Alice creates a contract where Deposit timeout would lead to Close 
\item Bob doesn't know or test the timeout path. Even if Bob is a programmer
- he might not be aware that Closing contract would cause collateral
refund 
\item Alice puts collateral and gets notional from Bob 
\item Alice doesn't pay for a loan (Deposit timeout) 
\item Alice gets her collateral back 
\item Bob loses notional 
\end{enumerate}
In order to prevent this from happening, an additional ''Auto-Refund''
security check was introduced as part of Static Analysis {[}TODO link{]}
tooling which notifies users about all ``Close'' constructs that
can lead to automatic refunds and encourages users to write explicit
logic for edge cases.

@@

Unlike current popular DeFi lending approaches (like Harvest Finance)
- Marlowe ACTUS-based implementation tends to rely on ``good old''
matching instead of pooling. While pooling is proven to be superior
to order-book based approaches when it comes to automated market makers,
using it for lending has been shown to be more susceptible to attacks
(for example ``\href{https://www.coindesk.com/harvest-finance-24m-attack-triggers-570m-bank-run-in-latest-defi-exploit}{flash loan}'').

@@

\section*{TODO}
